Towards Proof-Reconstruction of Problems in Classical Propositional Logic in Agda

### 2017-09-04 paper/main.tex
[ paper ] todo notes added.

@@

% \toctitle{Lecture Notes in Computer Science}
% \tocauthor{Authors' Instructions}
\maketitle

>>>>

% --------------------------------------------------------------------------
% TODO notes
% --------------------------------------------------------------------------

\usepackage{xargs}
\usepackage[pdftex,dvipsnames]{xcolor}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]
  {\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]
  {\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]
  {\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25
        ,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]
  {\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]
  {\todo[disable,#1]{#2}}

@@

% \toctitle{Lecture Notes in Computer Science}
% \tocauthor{Authors' Instructions}

\listoftodos[Notes]

% TODO commands!
% \unsure
% \change
% \info
% \improvement
% \thiswillnotshow

\subfile{TODO.tex}

\maketitle

### 2017-09-04 paper/sections/acknowledgements.tex
[ paper ] todo notes added.

We thank Joe Leslie-Hurd for all support with \verb!Metis! prover.
We also acknowledge the work carry out by Alejandro G\'omez-Londo\~no~
\cite{Gomez-Londono2015} that was used as the basis code for our TSTP parsing
module.
And last but not least, we gratefully acknowledge Andreas Abel and Chalmers
University of Goteborg, Sweden for inviting us to be part of the Agda
Implementors’ Meeting XXV where we presented part of this paper.

>>>>

We thank Joe Leslie-Hurd for all support with \verb!Metis! prover.
We also acknowledge the work carried out by Alejandro G\'omez-Londo\~no~
\cite{Gomez-Londono2015} used as the basis code for our TSTP parsing
module.
And last but not least, we thankfully acknowledge Andreas Abel and Chalmers
University of Goteborg, Sweden for inviting us to be part of the Agda
Implementors’ Meeting XXV where we presented part of this paper.

### 2017-09-04 paper/sections/introduction.tex
[ paper ] todo notes added.

Proof reconstruction is a hard labor since it depends on the integration
of two complex systems. On the one hand, we have the automatic theorem provers
(henceforth ATP) and their specification logic. These tools are usually
classified in at least one of the following categories. A SAT solver
(e.g., \verb!zChaff!~\cite{Moskewicz2001} and \verb!MiniSat!~\cite{Een2004})
to prove unsatisfiability of CNF formulas, a QBF solver
(e.g., \verb!GhostQ!~\cite{Klieber2014} and \verb!DepQBF!~\cite{Lonsing2017})
to prove satisfiability and invalidity of quantified Boolean formulas, a SMT
solver (e.g \verb!CVC4!~\cite{Barrett2011}, \verb!veriT!~\cite{bouton2009},
and \verb!Z3!~\cite{DeMoura2008}) to prove unsatisfiability of formulas
from first-order logic with theories, and a prover for validity of
formulas from first-order logic with equality
(e.g., \verb!E!~\cite{Schulz:AICOM-2002}, \verb!leanCoP!~\cite{Otten2008},
\verb!Metis!~\cite{hurd2003first}, \verb!SPASS! ~\cite{Weidenbach2009} and
\verb!Vampire!~\cite{Riazanov1999}), high-order logic (e.g., \verb!Leo-II!
\cite{Benzmuller2008} and \verb!Satallax!~\cite{Brown2012}) or intuitonistic
logic (e.g., \verb!ileanCoP!~\cite{Otten2008}, \verb!JProver!
\cite{Schmitt2001}, and \verb!Gandalf!~\cite{Tammet1997}), among others.
On the other hand, we have the proof checkers, interactive theorem provers
(henceforth ITP) or proof assistants (e.g., \verb!Agda!~\cite{agdateam},
\verb!Coq!~\cite{coqteam}, \verb!Isabelle!~\cite{paulson1994isabelle}, and
\verb!HOL4!~\cite{norrish2007hol}).
The ITP tools provide us the logic framework to check and validate the
reply of the ATPs since they allow us to define the formal language for
the problems like operators, logic variables, axioms, and theorems.

A proof reconstruction tool provides such an integration in one direction,
the bridge between ATPs to ITPs. This is mostly a translation of the reply
delivered by the prover into the formalism of the proof assistant.
Because the formalism of the source (the evidence generated by the ATP) is
not necessarily the same logic in the target, the reconstruction turns out
in a ``reverse engineering'' task. Then, reconstructing a proof involves an
in-depth understanding deep understanding of the algorithms in the ATP and
the specification logic in the ITP.

What we need from the ATP tools is a proof-object in a consistent format
to work with, that is, a full script describing step-by-step with
exhaustive details and without ambiguities, the table of the derivation
to get the actual proof. For problems in classical propositional logic
(henceforth CPL), from a list of at least forty\footnote{ATPs available
from the web service \texttt{SystemOnTPTP} of the TPTP World.} ATPs, just
a few provers were able to deliver proofs (e.g., \verb!CVC4!~
\cite{Barrett2011}, \verb!SPASS!, and
\verb!Waldmeister!~\cite{hillenbrand1997})
and a little bit less replied with a proof in a
file format like \verb!TSTP!~\cite{sutcliffe2004tstp} (e.g., \verb!E!, \verb!Metis!,
\verb!Vampire!, and \verb!Z3!), LFSC~\cite{Stump2008}, or the
SMT-LIB~\cite{Bohme2011} format.

>>>>

Proof reconstruction is a hard labor since it depends on the integration
of two complex systems. On the one hand, we have the automatic theorem provers
(henceforth ATP) and their specification logic. These tools are usually
classified in at least one of the following categories. A SAT solver
(e.g., \verb!zChaff!~\cite{Moskewicz2001} and \verb!MiniSat!~\cite{Een2004})
to prove unsatisfiability of CNF formulas, a QBF solver
(e.g., \verb!GhostQ!~\cite{Klieber2014} and \verb!DepQBF!~\cite{Lonsing2017})
to prove satisfiability and invalidity of quantified Boolean formulas, a SMT
solver (e.g \verb!CVC4!~\cite{Barrett2011}, \verb!veriT!~\cite{bouton2009},
and \verb!Z3!~\cite{DeMoura2008}) to prove unsatisfiability of formulas
from first-order logic with theories, and a prover for validity of
formulas from first-order logic with equality
(e.g., \verb!E!~\cite{Schulz:AICOM-2002}, \verb!leanCoP!~\cite{Otten2008},
\verb!Metis!~\cite{hurd2003first}, \verb!SPASS! ~\cite{Weidenbach2009} and
\verb!Vampire!~\cite{Riazanov1999}), high-order logic (e.g., \verb!Leo-II!
\cite{Benzmuller2008} and \verb!Satallax!~\cite{Brown2012}) or intuitionistic
logic (e.g., \verb!ileanCoP!~\cite{Otten2008}, \verb!JProver!
\cite{Schmitt2001}, and \verb!Gandalf!~\cite{Tammet1997}), among others.
On the other hand, we have the proof checkers, interactive theorem provers
(henceforth ITP) or proof assistants (e.g., \verb!Agda!~\cite{agdateam},
\verb!Coq!~\cite{coqteam}, \verb!Isabelle!~\cite{paulson1994isabelle}, and
\verb!HOL4!~\cite{norrish2007hol}).
The ITP tools provide us the logic framework to check and validate the
reply of the ATPs since they allow us to define the formal language for
the problems like operators, logic variables, axioms, and theorems.

A proof reconstruction tool provides such an integration in one direction,
is a translator of the reply delivered by the prover into the formalism of
the proof assistant. Because the formalism of the source (the evidence
generated by the ATP) is not necessarily the same logic in the target, the
reconstruction turns out in a ``reverse engineering'' task.
Then, reconstructing a proof involves an in-depth understanding of the
algorithms in the ATP and the specification logic in the ITP.
What we need from the ATP tools is a proof-object in a consistent format
to work with, that is, a full script describing step-by-step with
exhaustive details and without ambiguities, the table of the derivation
to get the actual proof. For problems in classical propositional logic
(henceforth CPL), from a list of at least forty\footnote{ATPs available
from the web service \texttt{SystemOnTPTP} of the TPTP World.} ATPs, just
a few provers were able to deliver proofs (e.g., \verb!CVC4!~
\cite{Barrett2011}, \verb!SPASS!, and
\verb!Waldmeister!~\cite{hillenbrand1997})
and a little bit less replied with a proof in a
file format like \verb!TSTP!~\cite{sutcliffe2004tstp} (e.g., \verb!E!, \verb!Metis!,
\verb!Vampire!, and \verb!Z3!), LFSC~\cite{Stump2008}, or the
SMT-LIB~\cite{Bohme2011} format.

### 2017-09-04 paper/sections/metis-language-and-proofs.tex
[ paper ] todo notes added.

\verb!Isabelle/HOL! integrated !Metis! as a macro-step reconstruction
tool for explaining proof steps (usually CNF goals) replied from other ATPs like
\verb!CVC4!, \verb!Vampire!, or \verb!Z3!.
In \cite{Farber2015}, \verb!Metis! was used to provide a tactic for
\verb!HOL Light!; this tactic challenges others like \verb!MESON! tactic or the
\verb!leanCoP! tactic that reconstructs proofs delivered by
\verb!leanCoP!~\cite{Farber2016}.

@@

The \verb!rule! stands for the inference name, the other fields are
auxiliary arguments or useful information to apply the inference, and
the list of parents nodes.
For instance, in Fig.~\ref{fig:metis-proof-tstp}, \verb!strip! is the name of one
of the inference and it has no arguments and only one parent node, \verb!goal!.

@@

\subsection{Proof Terms}
\label{ssec:proof-terms}

A proof-object delivered in a \verb!Metis!' proof encodes a natural
deduction proof. Its deduction system uses six simple inference rules
(see Fig.~\ref{fig:metis-inferences}) to attempt to prove conjectures
by refutation (see. e.g. Fig.~\ref{fig:metis-example}).

@@

These proofs are directed acyclic graphs, trees of refutations. Each node stands
for an application of an inference rule and the
leaves in the tree represent formulas in the given problem. Each node is
labeled with a name of the inference rule
(e.g. \verb!canonicalize! in Fig.~\ref{fig:metis-example}). Each
edge links a premise with one conclusion. The proof graphs have in their
root the conclusion $\bot$ since \verb!Metis! delivers proof by refutation.

% \begin{figure}
% \centering
% \begin{bprooftree}
%   \centering\tt
%   \AxiomC{}
%   \RightLabel{\scriptsize negate}
%   \UnaryInfC{$\neg p$}
%   \RightLabel{\scriptsize strip}
%   \UnaryInfC{$\neg p$}
%   \AxiomC{}
%   \RightLabel{\scriptsize axiom}
%   \UnaryInfC{$p$}
%   \RightLabel{\scriptsize canonicalize}
%   \UnaryInfC{$p$}
%   \RightLabel{\scriptsize simplify}
%   \BinaryInfC{$\bot$}
%   \RightLabel{\scriptsize canonicalize}
%   \UnaryInfC{$\bot$}
% \end{bprooftree}
% % \caption{The \verb!Metis!' refutation tree for $p \vdash p$ from the
% % derivation in Fig.~\ref{fig:metis-proof-tstp}}
% \label{fig:metis-example}
% \end{figure}

@@

Despite the six simple rules in the \verb!Metis!' logic
kernel (see~Fig. \ref{fig:metis-inferences}), we found in \verb!TSTP!
derivations for CPL problems other inference rules.
They are \verb!canonicalize!, \verb!conjunct!, \verb!negate!, \verb!simplify!,
\verb!strip! and \verb!resolve!. We briefly review each of these rules
following no order except maybe by their frequency in the \verb!TSTP! derivations.

@@

\textit{Clausification.} The \verb!clausify! rule transforms a
propositional formula in clausal normal form or CNF,
a conjunction of clauses. Where a clause is the disjunction of zero or
more literals and a literal is an atom (positive literal) or a negation of an
atom (negative literal). This conversion is not unique and \verb!Metis! performs
this transformation using its own criteria.

@@

\textit{Normalization.}
\verb!Metis! uses the \verb!canonicalize! rule to normalize a
propositional formula that comes from an axiom or a definition.
It often converts the formula in one of its normal form,
the conjunctive normal form,%as \verb!clausify! does
the negative normal form or the disjunctive normal form.
Posterior, this rule simplifies the formula with some definitions, applying
recursively on the structure's formula conjunctions and disjunctions simplifications
to remove tautologies
(see Fig.~\ref{fig:conjunctive-disjunctive-simplification} with a list
of these theorems. We had assumed the commutative property for conjunction and
disjunction connectives).

@@

\textit{Resolution.} The \verb!resolve! rule (see, for example,
Fig.~\ref{fig:resolve-inference}) is the resolution
theorem with two argument, the positive literal for resolution and two
formulas from their own derivations. The positive literal $p$ must occur in
the first formula of the input and the negative literal must occur in the second
formula of the input
(see the resolve rule in \verb!Metis!' logic core in
Fig.~\ref{fig:metis-inferences}).

@@

\textit{Splitting a conjunct.}
The \verb!conjunct! rule extracts from a conjunction one of its conjuncts, it is
a generalization of the projection rules for the conjunction connective as
we can see in the following \verb!TSTP! derivation.

@@

\textit{Simplification.} %abbreviate formulas
The \verb!simplify! rule is an $n$-ary rule that performs simplification
of definitions. This rule transverses a list of formulas by
applying different theorems (e.g.\emph{modus pones}, \emph{modus tollens},
or \emph{disjunctive syllogism}) including the list of theorems used by
\verb!canonicalize! in Fig.~\ref{fig:conjunctive-disjunctive-simplification}
and the resolution theorem of \verb!resolve! rule.
Moreover, many things happen inside this \verb!Metis!' procedure and
it is fairly complex to grasp it completely.

>>>>

\verb!Isabelle/HOL! integrated \verb!Metis! as a macro-step reconstruction
tool for explaining proof steps (usually CNF goals) replied from other ATPs like
\verb!CVC4!, \verb!Vampire!, or \verb!Z3!.
In \cite{Farber2015}, \verb!Metis! was used to provide a tactic for
\verb!HOL Light!; this tactic challenges others like \verb!MESON! tactic or the
\verb!leanCoP! tactic that reconstructs proofs delivered by
\verb!leanCoP!~\cite{Farber2016}.

@@

The \verb!rule! stands for the inference name; the other fields are
supporting arguments or useful information to apply the reasoning step, and
the list of parents nodes.
For instance, in Fig.~\ref{fig:metis-proof-tstp}, \verb!strip! is the name of one
of the inference with no arguments and only one parent node, the \verb!goal!.

@@

\subsection{Proofs}
\label{ssec:metis-proofs}

A proof-object delivered in a \verb!Metis!' proof encodes a natural
deduction proof. The deduction system uses six inference rules
\cite{hurd2003first} (see Fig.~\ref{fig:metis-inferences}) to attempt to
prove conjectures by refutation (see. e.g., the last conclusion
in the TSTP derivation Fig.~\ref{fig:metis-example} is \emph{falsium}).

@@

These proofs are directed acyclic graphs, trees of refutations. Each node stands
for an application of an inference rule and the
leaves in the tree represent formulas in the given problem. Each node is
labeled with a name of the inference rule
(e.g., \verb!canonicalize! in Fig.~\ref{fig:metis-example}). Each
edge links a premise with one conclusion. The proof graphs have at their
root the conclusion $\bot$ since \verb!Metis! delivers proof by refutation.

\begin{figure}
\centering
\begin{bprooftree}\tt
  \AxiomC{}
  \RightLabel{\scriptsize negate}
  \UnaryInfC{$\neg p$}
  \RightLabel{\scriptsize strip}
  \UnaryInfC{$\neg p$}
  \AxiomC{}
  \RightLabel{\scriptsize axiom}
  \UnaryInfC{$p$}
  \RightLabel{\scriptsize canonicalize}
  \UnaryInfC{$p$}
  \RightLabel{\scriptsize simplify}
  \BinaryInfC{$\bot$}
  \RightLabel{\scriptsize canonicalize}
  \UnaryInfC{$\bot$}
\end{bprooftree}
% \caption{The \verb!Metis!' refutation tree for $p \vdash p$ from the
% derivation in Fig.~\ref{fig:metis-proof-tstp}}
\label{fig:metis-example}
\end{figure}

@@

Despite the six rules in the \verb!Metis!' logic
kernel (see~Fig. \ref{fig:metis-inferences}), we observed in \verb!TSTP!
derivations for CPL problems other inference rules.
They are \verb!canonicalize!, \verb!conjunct!, \verb!negate!, \verb!simplify!,
\verb!strip! and \verb!resolve!. We briefly review each of these rules
following no order except maybe by their frequency in the \verb!TSTP! derivations.

@@

\textit{Clausification.} The \verb!clausify! rule transforms a
propositional formula in clausal normal form or CNF,
a conjunction of clauses. Where a clause is the disjunction of zero or
more literals and a literal is an atom (positive literal) or a negation of an
atom (negative literal). This conversion is not unique, and \verb!Metis! performs
this transformation using its criteria.

@@

\textit{Normalization.}
\verb!Metis! uses the \verb!canonicalize! rule to normalize a
propositional formula that comes from an axiom or a definition.
It often converts each expression in one of its normal form,
the conjunctive normal form, the negative normal form or the disjunctive normal form.
Posterior, this rule simplifies the formula with some definitions, applying
recursively on the structure's formula conjunctions and disjunctions simplifications
to remove tautologies
(see Fig.~\ref{fig:conjunctive-disjunctive-simplification} with a list
of these theorems. We had assumed the commutative property for conjunction and
disjunction connectives).

@@

\textit{Resolution.} The \verb!resolve! rule is the resolution
theorem showed in Fig.~\ref{fig:metis-inferences}. Its applications
needs the positive literal for resolution and two
derivations to deduce the \emph{resolvent}.
The positive literal $p$ must occur in the formula from the first derivation,
and the negative literal must occur in the formula from the second derivation
(this follows the same pattern in \verb!Metis!' logic core in
Fig.~\ref{fig:metis-inferences} for the resolve inference).

@@

\textit{Splitting a conjunct.}
The \verb!conjunct! rule extracts from a conjunction one of its conjuncts; it is
a generalization of the projection rules for the conjunction connective as
we can see in the following \verb!TSTP! derivation.

@@

\textit{Simplification.} %abbreviate formulas
The \verb!simplify! rule is a $n$-ary inference that performs simplification
of definitions. This rule transverses a list of formulas by
applying different theorems (e.g., \emph{modus pones}, \emph{modus tollens},
or \emph{disjunctive syllogism}) including the list of theorems used by
\verb!canonicalize! in Fig.~\ref{fig:conjunctive-disjunctive-simplification}
and the resolution theorem of \verb!resolve! rule.
Moreover, many things happen inside this \verb!Metis!' procedure, and
it is somewhat complicated to grasp it completely.

### 2017-09-04 paper/sections/proof-reconstruction.tex
[ paper ] todo notes added.

Our system (see Fig~\ref{fig:proof-reconstruction-overview}) translates a \verb!TSTP!
file into an \verb!Agda! proof by parsing it with a \verb!Haskell! tool.
This tool reconstructs the proof tree, removing redundancies and
cutting unnecessary steps (see for example repetitions in the derivation shown in
Fig~\ref{fig:metis-example}). The output is \verb!Agda! code.
Then, we use \verb!Agda! as a proof checker to type-check the proof.
If type-checking succeed, then our proof is verified. Else, we have a failure
proof due by three possible actors. First, a soundness bug in the ATP or a bug
in the output printing. Second, a bug in the translation process in our \verb!TSTP! parsing
module, the inner algorithms in the analysis of the proof tree, the pretty printer
module, or a bug in \verb!Haskell!. Lastly, a bug in our proof checker \verb!Agda!.

@@

\verb!Agda! is an interactive system for constructing proofs and programs,
based on Martin-L\"{o}f's type theory and extended with records, parametrised
modules, among other features.
One of the main strengths of \verb!Agda! is its support for writing proofs,
which we shall call \verb!Agda!'s proof engine and it consists of: support for
inductively defined types, including inductive families, and function
definitions using pattern matching on such types, normalisation during type-
checking, commands for refining proof terms, coverage checker and termination
checker.
The inductive approach for representing classical propositional logic
is better because we benefit from \verb!Agda!'s proof engine and its Unicode
support that allows us writing proofs similar as we find in math text books.

@@

A propositional formula is a string of indivisible propositional atoms,
logical connectives $\{\wedge, \vee, \Rightarrow, \Leftrightarrow, \neg\}$
and logic constants $\{\top, \bot\}$.
In \verb!Agda!, we define the formula
as an inductive type using the keyword \texttt{data} and including every
connective as a constructor. We represent propositional formulas with
\verb!Prop! data type using the following definition.

>>>>

Our system (see Fig~\ref{fig:proof-reconstruction-overview}) translates a \verb!TSTP!
file into an \verb!Agda! proof by parsing it with a \verb!Haskell! tool.
This tool reconstructs the proof tree, removing redundancies and
cutting unnecessary steps (see for example repetitions in the derivation shown in
Fig~\ref{fig:metis-example}). The output is \verb!Agda! code.
Then, we use \verb!Agda! as a proof checker to type-check the proof.
If type-checking succeed, then our proof is checked. Else, we have a failure
proof due to three possible factors. First, a soundness bug in the ATP or a bug
in the output printing. Second, a bug in the translation process in our \verb!TSTP! parsing
module, the inner algorithms in the analysis of the proof tree, the pretty printer
module, or a bug in \verb!Haskell!. Lastly, a bug in our proof checker \verb!Agda!.

@@

\verb!Agda! is an interactive system for constructing proofs and programs,
based on Martin-L\"{o}f's type theory and extended with records, parametrised
modules, among other features.
One of the main strengths of \verb!Agda! is its support for writing proofs,
which we shall call \verb!Agda!'s proof engine. It consists of: support for
inductively defined types, including inductive families, and function
definitions using pattern matching on such types, normalisation during type-
checking, commands for refining proof terms, coverage checker and termination
checker.
The inductive approach for representing classical propositional logic
is better because we benefit from \verb!Agda!'s proof engine and its Unicode
support that allows us writing proofs similar as we find in math text books.

@@

A propositional formula is a experssion of indivisible propositional atoms,
logical connectives $\{\wedge, \vee, \Rightarrow, \Leftrightarrow, \neg\}$
and logic constants $\{\top, \bot\}$.
In \verb!Agda!, we define the formula
as an inductive type using the keyword \texttt{data} and including every
connective as a constructor. We represent propositional formulas with
\verb!Prop! data type using the following definition.

### 2017-09-04 paper/sections/related-work.tex
[ paper ] todo notes added.

\verb!Sledgehammer! is a tool for \verb!Isabelle! proof assistant
~\cite{paulson1994isabelle} that provides a full integration of automatic
theorem provers including ATPs
(see, for example, \cite{meng2006automation}, \cite{blanchette2013extending}
and \cite{Fleury2014}) and SMT solvers (see, for example,
\cite{hurlin07practical}, \cite{bohme2010},
\cite{blanchette2013extending}, and \cite{Fleury2014}) with
\verb!Isabelle/HOL! \cite{nipkow2002isabelle}, the specialization of
\verb!Isabelle! for higher-order logic. Sultana, Benzm{\"{u}}ller, and
Paulson~\cite{Een2004} integrates \verb!Leo-II! and \verb!Satallax!, two
theorem provers for high-order logic with \verb!Isabelle/HOL! proposing a
modular proof reconstruction work flow.

@@

Hurd~\cite{Hurd1999} integrates the first-order resolution prover
\verb!Gandalf! with \verb!HOL!~\cite{norrish2007hol}, a high-order theorem
prover, following a LCF model implementing the tactic \verb!GANDALF_TAC!.
For \verb!HOL Light!, a version of \verb!HOL! but with a simpler logic core,
the SMT solver \verb!CVC4! was integrated and Kaliszyk and
Urban~\cite{kaliszyk2013} reconstruct proofs from different ATPs with their
\verb!PRocH! tool, replaying the detailed inference steps from the ATPs with
internal inference methods implemented in \verb!HOL Light!.

@@

\verb!Waldmeister! is an automatic theorem prover for unit equational
logic \cite{hillenbrand1997}.
Foster and Struth~\cite{foster2011integrating} integrate \verb!Waldmeister! into
\verb!Agda!~\cite{agdateam}. This integration requires a proof
reconstruction step but authors' approach is restricted to pure
equational logic --also called identity theory~\cite{humberstone2011}--
that is, first-order logic with equality but no other predicate symbols
and no functions symbols~\cite{appel1959}.

>>>>

\verb!Sledgehammer! is a tool for \verb!Isabelle! proof assistant
~\cite{paulson1994isabelle} that provides a full integration of automatic
theorem provers including ATPs
(see, for example, \cite{meng2006automation}, \cite{blanchette2013extending}
and \cite{Fleury2014}) and SMT solvers (see, for instance,
\cite{hurlin07practical}, \cite{bohme2010},
\cite{blanchette2013extending}, and \cite{Fleury2014}) with
\verb!Isabelle/HOL! \cite{nipkow2002isabelle}, the specialization of
\verb!Isabelle! for higher-order logic. Sultana, Benzm{\"{u}}ller, and
Paulson~\cite{Een2004} integrates \verb!Leo-II! and \verb!Satallax!, two
theorem provers for high-order logic with \verb!Isabelle/HOL! proposing a
modular proof reconstruction work flow.

@@

Hurd~\cite{Hurd1999} integrates the first-order resolution prover
\verb!Gandalf! with \verb!HOL!~\cite{norrish2007hol}, a high-order theorem
prover, following an LCF model implementing the tactic \verb!GANDALF_TAC!.
For \verb!HOL Light!, a version of \verb!HOL! but with a simpler logic core,
the SMT solver \verb!CVC4! was integrated, and Kaliszyk and
Urban~\cite{kaliszyk2013} reconstruct proofs from different ATPs with their
\verb!PRocH! tool, replaying the detailed inference steps from the ATPs with
internal inference methods implemented in \verb!HOL Light!.

@@

\verb!Waldmeister! is an automatic theorem prover for unit equational
logic \cite{hillenbrand1997}.
Foster and Struth~\cite{foster2011integrating} integrate \verb!Waldmeister! into
\verb!Agda!~\cite{agdateam}. This integration requires a proof
reconstruction step, but authors' approach is restricted to pure
equational logic --also called identity theory~\cite{humberstone2011}--
that is, first-order logic with equality but no other predicate symbols
and no functions symbols~\cite{appel1959}.
